---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(rpart)
library(randomForest)
library(wakefield)
library(dplyr)

A.sim <- function(matrix.pi) {
  N <- nrow(matrix.pi) # sample size
  K <- ncol(matrix.pi) # treatment options
  if (N<=1 | K<=1) stop("Sample size or treatment options are insufficient!")
  if (min(matrix.pi)<0) stop("Treatment probabilities should not be negative!")
  
  # normalize probabilities to add up to 1 and simulate treatment A for each row
  probs <- t(apply(matrix.pi,1,function(x){x/sum(x,na.rm = TRUE)}))
  A <- apply(probs, 1, function(x) sample(0:(K-1), 1, prob = x))
  return(A)
}
```

# clusterQ

<!-- badges: start -->
<!-- badges: end -->

A clustered Q-learning algorithm with M-out-of-N cluster bootstrap for making inference on tailoring variables for optimal dynamic treatment regimes (DTR) from clustered sequential multiple assignment randomized trials (clustered SMART). This tool is developed based on paper by Speth, et al. (2024).


## Installation

You can install the development version of MOTRL from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("SelinaSong0412/clusterQ")
```

## Example

See below for two examples of using MOTRL to estimating tolerant DTR.

```{r example}
library(clusterQ)
```

**(a). **

Here we simulate a clustered SMART data. 

